---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output: html_notebook
---
# Lets load in any necessary packages
```{r}
## I may not use all of these packages
library(tidyverse)
library(dplyr)
library(readxl)
library(janitor)
library(here)
library(lubridate)
library(stringr)
```

```{r}
# test where the top level of the project directory is
here::here()

# loading packages
library(janitor)

# loading in the cleaned data
cleaned_halloween_candy_data <- read_csv(here("clean_data/cleaned_halloween_candy_data_pivot_long.csv"))
view(cleaned_halloween_candy_data)
glimpse(cleaned_halloween_candy_data)

# having done a fair bit of work on country already, i've decided to add the following at the start of the process ie set everything in the country column to lower case - this should reduce the amount of tidying required.

cleaned_halloween_candy_data$country <- tolower(cleaned_halloween_candy_data$country)
```
# now to have a look at the data
```{r}
glimpse(cleaned_halloween_candy_data)
view(cleaned_halloween_candy_data)
```

# let's see if there are any columns that we can remove due to having no data - nope, all columns have at least some data.
```{r}
cleaned_halloween_candy_data %>%
  summarise(across(.fns = ~ sum(is.na(.x))))
```

# Now to have a look at some of the data in specific columns
```{r}

HC_country_overview <- cleaned_halloween_candy_data %>% 
  group_by(country) %>% 
  summarise(character_count = n()) %>% 
  arrange(desc(character_count))
head(HC_country_overview, 10)

# use
HC_data_cleaning_v2_long <- cleaned_halloween_candy_data %>% 
  group_by(country) %>% 
  distinct(country) %>% 
  summarise(character_count = n()) %>% 
  arrange(character_count)

view(HC_data_cleaning_v2_long)
# straight away we can see that headings related to the USA need unified - so I need to hand back to data cleaning.

```

```{r}
test <- HC_data_cleaning_v2_long %>% 
  mutate(country = str_replace_all(country, c("'merica" = "usa", "^a$" = "unknown", "a tropical island south of the equator" = "unknown", "(^[uni][a-z]+[:space:]+[s][a-t]{4})" = "usa", "([")))

# works to a point
test <- HC_data_cleaning_v2_long %>% 
  mutate(country = str_replace_all(country, c("atlantis" = "unknown","california" = "usa", "canada`" = "canada","^can$" = "canada","canae" = "canada","cascadia" = "unknown","denial" = "unknown","earth" = "unknown","endland" = "england","españa" = "spain","eua" = "uae","fear and loathing" = "unknown","god's country" = "unknown","i don't know anymore" = "unknown","i pretend to be from canada, but i am really from the united states." = "usa","insanity lately" = "unknown", "'merica" = "usa","merica" = "usa","murica" = "usa","murrika" = "usa","n. ausa" = "usa","narnia" = "unknown","neverland" = "unknown","new jersey" = "usa","new york" = "usa","north carolina" = "usa","not the usa or canada" = "usa","one of the best ones" = "unknown","pittsburgh" = "usa","see above" = "unknown","somewhere" = "unknown","soviet canuckistan" = " unknown","sub-canadian north ausa... usa" = "usa","subscribe to dm4uz3 on youtube" = "unknown","the best one - usa" = "usa","the netherlands" = "netherlands","the republic of unknown" = "unknown","the united states" = "usa", "america" = "usa", "usaa" = "usa","ausa" = "usa","the yoo ess of aaayyyyyy" = "usa","there isn't one for old men" = "unknown","this one" = "unknown","trumpistan" = "usa","u s" = "usa","^u.k.$" = "uk","^u.s.$" = "usa","^u.s.a.$" = "usa","unhinged states" = "usa","^unied states$" = "usa","unite states" = "usa","^ud$" = "unknown", "ahem....amerca" = "usa","^united states of usa$" = "usa")))

test2 <- test %>% 
  mutate(country = str_replace_all(country, c("united states of usa" = "usa", "^united kingdom" = "uk", "united staes" = "usa", "^united sates" = "usa", "^united state%" = "usa")))


test %>% 
  group_by(country) %>% 
  distinct(country) %>% 
  summarise(character_count = n()) %>% 
  arrange(character_count)

```




```{r}
cleaned_halloween_candy_data %>%
  summarise(across(.fns = ~ sum(is.na(.x))))
```

```{r}
joined_BD_record_ID_and_SD_record_ID_IJ %>%
  summarise(across(.fns = ~ sum(is.na(.x))))
```

```{r}
# to check that latitude and longitude data don't have any unusual data.

library(assertr)
assertr_long_lat_values_in_seabird_data_check <- function(seabird_data){

  seabird_data %>% 
    verify(LAT >= -90 & LAT <= 90) %>% 
    verify(LONG >= -180 & LONG <= 180)
}
assertr_long_lat_values_in_seabird_data_check(cleaned_seabird_data)

# having run this test there are a number fails due to missing data (eg NA's) - will remove the NA's in the data_cleaning_script (see line 100 in the file 'data_cleaning_script.R')
# as an observation, i wonder whether I could fill the data using over observations eg, same person on the same date - therefore I know where they were (however I don't know that I have time at the present).

```

```{r}
# picking back up the data after changing lat and long NA's to 0's

cleaned_seabird_data_lat_long_NAs_replaced <- read_csv(here("clean_data/cleaned_seabird_data_lat_long_zeros.csv"))
view(cleaned_seabird_data_lat_long_NAs_replaced)
```


```{r}
# rerunning our assertr test to check that our missing data (NA's) have been replaced by 0
assertr_long_lat_values_in_seabird_data_check <- function(seabird_data){

  seabird_data %>% 
    verify(LAT >= -90 & LAT <= 90) %>% 
    verify(LONG >= -180 & LONG <= 180)
}
assertr_long_lat_values_in_seabird_data_check(cleaned_seabird_data_lat_long_NAs_replaced)

# no errors this time round.
```
# now to clean the names
```{r}
library(janitor)
cleaned_names_cleaned_seabird_data <- clean_names(cleaned_seabird_data_lat_long_NAs_replaced)
names(cleaned_names_cleaned_seabird_data)

```

# Q1: Which bird had the most individual sightings?
```{r}
summarised_scientific_name <- cleaned_names_cleaned_seabird_data %>% 
  group_by(species_common_name_taxon_age_sex_plumage_phase, species_scientific_name_taxon_age_sex_plumage_phase, species_abbreviation) %>% 
summarise(character_count = n()) %>% 
arrange(desc(character_count))
head(summarised_scientific_name)

```

# Q2: Which bird had the highest total count?
```{r}

cleaned_names_cleaned_seabird_data %>% 
  group_by(species_common_name_taxon_age_sex_plumage_phase, species_scientific_name_taxon_age_sex_plumage_phase, species_abbreviation, count) %>% 
  summarise(count = sum(count)) %>% 
  arrange(desc(count))

# therefore bird with highest total count = short tailed shearwater

# to note I tried a number of alternatives eg I figured that that column name with the most accurate classification of the birds would be the Species Abbreviation however despite repeated attempts I could not get it down to one count per entry:

data_new2 <- as.data.frame(apply(test, 2, function(x) gsub("\\s+", "", x))) # Removes blanks (which is ok as column seems to contain only 1 word entries - this is an attempt to consolidate multiple entries into 1), however changes count column to character; I address this in the next step.
glimpse(data_new2)

data_new2 %>% 
  mutate(count = as.numeric(count)) %>% # changing count column back to a double (sourced from the internet)
  group_by(species_common_name_taxon_age_sex_plumage_phase, species_scientific_name_taxon_age_sex_plumage_phase, species_abbreviation, count) %>% 
  summarise(count = sum(count)) %>% 
  arrange(desc(count))

test <- summarised_count_of_sightings %>% 
  group_by(species_abbreviation, count) %>% 
    summarise(count = sum(count)) %>% 
    # sum(count, NA, na.rm = TRUE) %>% 
  arrange(desc(count))

view(test)

# just using this to veiw column names
names(cleaned_names_cleaned_seabird_data)

##### ignore these - various unsuccessful attempts

test3 <- read_csv(here("clean_data/test_specific_columns.csv"))
test3 <- clean_names(test3)
data_new3 <- as.data.frame(apply(test3, 2, function(x) gsub("\\s+", "", x))) # Remove blanks
glimpse(data_new3)

data_new3 %>% 
  mutate(count = as.numeric(count)) %>%
  group_by(species_abbreviation, count) %>% 
  summarise(count = sum(count)) %>% 
  arrange(desc(count))
  
test2 <- test %>% 
  group_by(species_abbreviation, count) %>% 
  summarise(count = sum(count)) %>% 
  arrange(desc(count))
  head(test2)

summarised_count_of_sightings <- cleaned_names_cleaned_seabird_data %>%
  group_by(species_common_name_taxon_age_sex_plumage_phase, species_scientific_name_taxon_age_sex_plumage_phase, species_abbreviation, count) %>% 
  summarise(count = sum(!is.na(count))) %>% 
arrange(desc(count))
head(summarised_count_of_sightings)

summarised_count_of_sightings <- cleaned_names_cleaned_seabird_data %>%
  select(species_abbreviation, count) %>% 
  arrange(desc(count))
view(summarised_count_of_sightings)


```

```{r}
# alsternative approach provided by a colleague - I'm personally still unclear as to whether I'm just misunderstanding the quesion.

# group by bird name and take a sum of count, join this back into the table
# to allow more names info to be pulled, arrange and slice head to get top hit
cleaned_names_cleaned_seabird_data %>% 
  group_by(species_common_name_taxon_age_sex_plumage_phase) %>% 
  summarise(total_count = sum(count)) %>% 
  arrange(desc(total_count))

cleaned_names_cleaned_seabird_data %>% 
  group_by(species_common_name_taxon_age_sex_plumage_phase) %>% 
  summarise(total_count = sum(count)) %>% 
  left_join(cleaned_names_cleaned_seabird_data, "species_common_name_taxon_age_sex_plumage_phase") %>% 
  select(species_common_name_taxon_age_sex_plumage_phase, 
         species_scientific_name_taxon_age_sex_plumage_phase, 
         species_abbreviationn, total_count) %>%
  arrange(desc(total_count)) %>% 
  slice_head()
  
# Answer: A Black-billed gull, count = 7732
```

# Q3 Which bird had the highest total count above a latitude of -30?
```{r}
cleaned_names_cleaned_seabird_data %>% 
  group_by(species_common_name_taxon_age_sex_plumage_phase, species_scientific_name_taxon_age_sex_plumage_phase, species_abbreviation, lat, count) %>% 
  filter(lat > -30) %>% 
  summarise(count = sum(count)) %>% 
  arrange(desc(count))
```

# Q4: How many different types of birds were only ever seen in groups of 1?
```{r}
cleaned_names_cleaned_seabird_data %>% 
  group_by(species_common_name_taxon_age_sex_plumage_phase, species_scientific_name_taxon_age_sex_plumage_phase, species_abbreviation, count) %>% 
  filter(count == 1) %>% 
  summarise(character_count = n()) %>% 
  arrange(desc(character_count))

# Answer = 298
```

# Q5: How many penguins were seen?
```{r}
library(stringr)
cleaned_names_cleaned_seabird_data %>% 
  filter(str_detect(species_common_name_taxon_age_sex_plumage_phase, pattern = "[Pp]enguin")) %>%
  group_by(species_common_name_taxon_age_sex_plumage_phase, species_scientific_name_taxon_age_sex_plumage_phase, species_abbreviation, count) %>% 
  summarise(character_count = n()) %>% 
  arrange(desc(character_count))


```


